\plaintitle

### 2016-07-16 cscw/papers/2017/labor/introduction.tex
cleanup
cleanup?
some other intro revisions..

\documentclass[trackingWork]{subfiles}
\makeatletter
\def\blx@maxline{77}
\makeatother
\begin{document}
\section{Introduction}
% what would a cynical reader make bullet points out of?
% each sentence feels like a thought that's unrelated to the next one
The past decade has seen a proliferation of transient labor.
% beginning with ``information work'' such as on
% Amazon Mechanical Turk (AMT).
% ??? ->
Stemming from the recognition that
complex tasks might be achievable
--- and indeed trivial ---
through the effective management of the ``crowd'',
researchers have rapidly begun exploring
the space for opportunities to facilitate new (and revisit familiar) forms of work
\cite{bernsteinSoylent,redballoon}.
%  ??? -->
Broadly, this work has imagined the various applications of ``gig''--based work,
and making it work for ``requesters''
--- academics, industries, and to some extent end--users
who might benefit from the product of crowd--powered work
\cite{hong2015group,jonBrelig,paolacci2010running}.
% ??? -->
Since then,
the application of digitally mediated environments acting as rapid ``hiring halls''
for on--demand tasks
has found purchase far beyond the context of ``information work'' like Amazon Mechanical Turk (AMT);
livery services (Uber), cleaning (Handy), and various other services (TaskRabbit)
% parentheticals giving examples?
have garnered the interest of both industry and research.

>>>>

% %!TEX root = trackingWork.tex
\documentclass[trackingWork]{subfiles}
\makeatletter
\def\blx@maxline{77}
\makeatother
\begin{document}
\section{Introduction}
% each sentence feels like a thought that's unrelated to the next one
In the past 10 years, TEST TEST TEST
Amazon Mechanical Turk (AMT) has been the site of a flourishing economy for ``information work''
\cite{paolacci2010running}.

The past decade has seen a proliferation of a new type of labor measured in discrete, atomic tasks and on--off gigs;
the workers that do these tasks float into and out of the labor market so freely that it has been called ``transient''
by some
\onlyinsubfile{\ali{who talks about this? definitely \cite{solinger1995chinese} but not the right context! (although, super interesting)}}.
Stemming from the recognition that
complex tasks might be achievable
--- and indeed trivial ---
through the effective management of the ``crowd'',
researchers have rapidly begun exploring
the space for opportunities to facilitate new (and revisit familiar) forms of work
\cite{bernsteinSoylent,redballoon}.
%  ??? -->
Broadly, this work has imagined the various applications of ``gig''--based work,
and making it work for ``requesters''
--- academics, industries, and to some extent end--users
who might benefit from the product of crowd--powered work
\cite{hong2015group,jonBrelig,paolacci2010running}.
% ??? -->
Since then,
the application of digitally mediated environments acting as rapid ``hiring halls''
for on--demand tasks
has found purchase far beyond the context of ``information work'' like Amazon Mechanical Turk (AMT);
livery services (Uber), cleaning (Handy), and various other services (TaskRabbit)
% parentheticals giving examples?
have garnered the interest of both industry and research.

### 2016-07-16 cscw/papers/2017/labor/trackingWork.tex
some other intro revisions..

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}
\definecolor{PineGreen}{HTML}{008800}
\definecolor{BrickRed}{HTML}{FF0000}
\newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
\newcommand{\ali}[1]{{\color{BrickRed}[\itshape al2: #1\upshape]}}

@@

\keywords{\plainkeywords}

\section{Introduction}
The past decade has seen
paid microtasking, the ``sharing economy'', and other instantiations of on--demand contract work
grow to occupy the fascination of both academic circles and our culture as a whole
\cite{Ross,CrowdsourcingUserStudies,scholz2012digital}.
The research community has made connections between this emergent form of work
and the historically situated ``piecework'',
as well as a number of observations relating to
the frustration workers feel stemming from the management of this work,
but situating the causes of these frustrations either in the larger body of scholarship on crowd work,
or crowd work itself,
has proven difficult;
the connection between crowd work and a historical parallel contextualizing its ongoing developments 
hasn't been deeply made yet
\cite{turkopticon,dynamo,uberAlgorithm}.

While much of this work appears to gesture toward
the parallels between contemporary on--demand work and piecework,
it's proven difficult to bring the totality of these observations into focus using one theoretical lens.
This paper will attempt to do so by arguing that
the topics social computing researchers have investigated are
not just parallel to historical piecework and
the process of factorization that took place in the early \nth{20} century;
indeed, these phenomena \textit{precipitate and reinforce each other}.

In the reflection on the literature published in the last 5 years since
\citeauthor{crowdworkFuture}'s 
\citetitle{crowdworkFuture},
we notice a broader trend describing the change in work that's being done
\cite{crowdworkFuture}.
Following the improvements in
breaking down tasks,
delegating work, and
managing workers more broadly,
crowd--powered work has continued to parallel historical piecework's trajectory
by outgrowing ad--hoc worker groups and coalescing into more formal working groups,
for example through the use of a ``white list'' of familiar workers.
We might call this the regularization of factory work.

Making sense of the broader field of crowd work
by looking at the movements toward
distributing work,
routinizing and breaking down tasks,
and externally managing workers
as linked to one another, 
we give ourselves a framing of contemporary piecework that explains,
and arguably predicts,
what we have seen thus far
--- and perhaps what we should expect to see going forward.

This paper will attempt to demonstrate that the familiarity
that we incidentally observe between what we generally call crowd work \& gig work
and the historical practice of piecework and subsequent factory work
is more than passing;
that the phenomena researchers have observed in on--demand digitally mediated labor markets
were inevitable milestones in the birth and life of piecework.
We will set out to show that this perspective of crowd work as an embodiment of piecework
predicts the myriad outcomes of contemporary crowd work, including
our developments of work--flows further abstracting work,
the troubling effects of those developments on factors such as pay and work quality, and
many if not all of the frustrations researchers have discovered among workers
in the study of crowd and gig work.

``But'',
as \citeauthor{scholz2012digital} points out,
``it would be wrong to conclude that
in the realm of digital labor there is nothing new under the sun''
\cite{scholz2012digital}.
For one, information work
(e.g. Amazon Mechanical Turk (AMT) \& Upwork)
and digitally managed work in general make it substantially easier
to keep workers geographically dispersed
compared to the factory workers that followed home--based piecework.
This aspect of work renders workers virtually invisible to the task solicitors
--- or ``requesters'' ---
despite being significantly more connected than historical pieceworkers were % this is a weak point...
\cite{turkopticon}.
Secondly, while
factory foremen and other middle--men
traditionally mediated the relationships between workers and managers,
today the visible agents are \textit{systems};
rather than employing individuals who ``personified the functions of management''
and can thus negotiate workers' needs,
socio--technical systems mediate these interactions % often with unrewarding outcomes
\cite{wray1949marginal}.
% \cite{storiesIraniSilberman}.
Finally, this characteristic of modern piecework makes
worker coordination for collective advocacy and action significantly more difficult,
necessitating special consideration to maintain the inertia of collective will
while focusing that energy productively
\cite{dynamo}.
% \ali{I really don't like this paragraph\dots}
% here are undoubtedly more examples of differences.
% How do I not stake a claim that this is comprehensive?
% Or should I really flesh this out and make it (comprehensive)?

\subsection{Piecework as a lens to understand gig work}

\citeauthor{crowdworkFuture}
write of the future of crowd work in an effort to investigate
whether crowd work will be a worthwhile form of work or another instantiation of ``piecework''
\cite{crowdworkFuture}.
``Piecework'' here makes use of
a term historically used to describe work done in the home,
in manageable tasks,
often involving clear instructions
and payment only for work completed, not work done
(the differentiation, here, being that
one would be paid for the \textit{output} of the work,
not the \textit{duration}).
% \msb{split up that sentence. One is the def'n of piecework, the other is the comparison you're drawing}
Given the scope,
we can frame piecework and on--demand labor
% \msb{
% \sout{wait, are you using gig work as the umbrella term? I think you need a category term that covers all of them}
% \ali{let's use ``on--demand labor'', since
% that seems general enough without accidentally including historical piecework}}
as sharing these important similarities:
\begin{inlinelist}
\item this form of work began in the home
\item the worker is paid for each discrete piece of work done, regardless of time or effort; and
\item the worker's status
(not only socially, but also economically)
is ambiguous, or at least the subject of some controversy.
\end{inlinelist}.

% \subsection{Context}
In the past decade, researchers have observed frustration
grow among on--demand workers,
with expression of this frustration spanning a wide range of tactics
\cite{uberAlgorithm,turkopticon,dynamo}.
Attempting to make sense of these case studies has been challenging
in part because
a wholly encompassing framework for understanding this form of work
has thus proven difficult to capture.

This paper attempts to make sense of the broader research on this new form of work,
\textit{on--demand labor}, by evaluating this work through a more familiar lens: piecework.
More concretely, by looking at task--based or ``gig'' work as
an instantiation and continuation of piecework,
and by looking for patterns of behavior that the corresponding literature predicts
on this basis, we can do the following:
\begin{inlinelist}
  \item we can make sense of the phenomena so far as part of a much larger series of interrelated events;
  \item we can bring into focus the ongoing work among workers, system--designers, and researchers in this space; and finally,
  \item we can offer predictions of what social computing researchers, and workers themselves, should expect to see on the horizon of on--demand work.
\end{inlinelist}

We'll look at a broad range of cases under a number of major themes
we propose as broadly describing the types of research being done in crowd work
and more generally in what we argue is contemporary piecework.
After validating this lens as a way of reasoning about on--demand labor,
we'll attempt to use this perspective to suggest areas of research worth anticipating,
and developments we should expect to see in the maturation of digitally mediated work.

>>>>

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}
\definecolor{PineGreen}{HTML}{008800}
\definecolor{BrickRed}{HTML}{FF0000}
\newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
\newcommand{\ali}[1]{{\color{BrickRed}[al2: #1]}}

@@

\keywords{\plainkeywords}
\subfile{introduction.tex}
